---
title: "Boostrapping mortality rates"
author: Daniel Zuleta
date: July, 2023 ATFS-MWG
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = F}
knitr::opts_chunk$set(warning = F, message = F)
```

### 0. Simulated dataset ###
We will use a simulated dataset with the necessary metrics at the quadrat level for each period (note that it can be done at any scale)
```{r}


# Set  the seed to generate the same simulation
set.seed(23842232)

# Generate the data
metrics.period <- data.frame(
  quadrat = LETTERS[1:25],
  N0 = sample(100:200, 25, replace = T),
  time.diff = sample(2:5, 25, replace = T)
)
for(i in 1:nrow(metrics.period)) metrics.period$NsT[i] <-  floor(metrics.period$N0[i]*sample(seq(0.85, 1, 0.1), 1))

```
Simulated dataset:
N0: Initial number of individual trees at t = 0 in a given quadrat; time.diff: time census interval in years for a given quadrat; NsT: Number of surviving trees at t = T in a given quadrat
```{r}
metrics.period
```

## 1. Estimate the true mortality rates ##
We will use equation 5 in Kohyama et al. (2018) DOI:	10.1111/2041-210X.12929
```{r}
# Annual mortality rates Kohyama et al. 2018 (Table 1).
true.rates.per.quadrat <- 1 - (metrics.period$NsT / metrics.period$N0)^(1/metrics.period$time.diff)
true.rates.plot <- 1 - (sum(metrics.period$NsT) / sum(metrics.period$N0))^(1/mean(metrics.period$time.diff))

true.rates.per.quadrat # true mortality at the stand level
true.rates.plot # true mortality at the quadrat level
```

## 2. Bootstrap over quadrats to get the confidence intervals ####

```{r}
# define the number of simulations for the bootstrap (usually 1000)
nsim = 1000
# empty vector
m.boot = NULL

for (i in 1:nsim){
  temp <- 1:nrow(metrics.period)
  temp2 <- sample(temp, replace = T)
  
  metrics.period.i <- metrics.period[temp2,]
  rates.i <- 1 - (metrics.period.i$NsT / metrics.period.i$N0)^(1/metrics.period.i$time.diff) # Here you should do all the calculations in case you want to bootstrap more variables e.g., biomass
  
  # estimating the average mortality from the realization i
  m.i <- mean(rates.i)
  
  # SAVE main results from this iteration
  m.boot = c(m.boot, m.i)
}
```

## 3. Get the confidence intervals

```{r}
m.lo <- quantile(m.boot, probs = c(0.025))
m.hi <- quantile(m.boot, probs = c(0.975))

true.rates.plot
m.lo
m.hi

```



